---
title: "WGCNA Demo"
author: 'Shaurya Jauhari (Email: shauryajauhari@gzhmu.edu.cn)'
date: '`r paste(Sys.Date())`'
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Installing the package and setting up the options.

```{r package installation}

install.packages("BiocManager", 
                 repos='http://cran.us.r-project.org',
                 dependencies = TRUE)
BiocManager::install("WGCNA")

install.packages("ggdendro",
                 repos='http://cran.us.r-project.org',
                 dependencies = TRUE)

## Setting options

options(stringsAsFactors = FALSE)

#enableWGCNAThreads() ## Enabling multi-threads in processing.

library(WGCNA)
library(ggdendro)
library(ggplot2)
library(cowplot)

```

Importing data files from female and male liver tissues from mice, and exploring
them.

```{r data import}

mydataf <- read.csv("../FemaleLiver-Data/LiverFemale3600.csv", header = TRUE) 
colnames(mydataf)

head(mydataf)

## LocusLinkID and ProteomeID are annotations from the said databases
## http://www.ncbi.nlm.nih.gov/LocusLink/

```

Moving on, we extract expression data from the master dataframe. Recall that the rows represent genes and the columns represent different samples (mice) in the original data. WGCNA requires that genes be given in columns.

```{r data cleaning}

exprdata = as.data.frame(t(mydataf[, -c(1:8)]))
names(exprdata) = mydataf$gene_symbol
rownames(exprdata) = names(mydataf)[-c(1:8)]


## Let us consider a subset of data for this demonstration. We'll use first 500 features.

exprdata <- exprdata[,1:500]
```

```{r missing values and outliers}

gsg = goodSamplesGenes(exprdata, verbose = 3)
gsg$allOK

```
## Scale Free Topology

A scale free network topology is the one where all nodes's degree distribution is in abidance to power law ,i.e. **P(k) ~ k^- beta)**. If any nodes have to be added to this connected network, the degrees are accordingly adjusted. 

```{r Soft_Thresholding}

trial_powers <- c(c(1:10), seq(from=12, to=20, by=2))
sft_thresh <- pickSoftThreshold (exprdata,
                                dataIsExpr = TRUE,
                                powerVector = trial_powers,
                                corFnc = cor,
                                corOptions = list(use = 'p'),
                                networkType = "unsigned")
cat("The best estimate for use is", sft_thresh$powerEstimate, ".")
print(sft_thresh$fitIndices)

```

It has been determined that *beta=6 is the best exponent for unsigned networks, and beta=12 is the optimal one for signed networks*.

Typically, the degrees per node will "steadily" decrease with the increase in the number of nodes of the network cluster. This is another way of understanding the notion of scale-free topology.  


```{r Plotting the estimated powers against slopes}

# Scale-free topology fit index as a function of the soft-thresholding power

p1 <- ggplot(data = as.data.frame(sft_thresh),
       aes (x = sft_thresh$fitIndices[,1],
     y = (-sign(sft_thresh$fitIndices[,3]) * sft_thresh$fitIndices[,2]),
     label = trial_powers)) + 
  geom_label()+
  
  # the sign expression returns the vector of signs of estimated slopes, multiplied by the correlation coefficient.

     labs( x="Soft Threshold (power)",
     y="Scale Free Topology Model Fit, signed R^2", title = "Scale Independence", tag = "A")+
  theme_classic()+
  geom_vline(xintercept = 8, color="red") # Red line corresponds to using an R^2 cut-off


# Mean connectivity as a function of the soft-thresholding power

p2 <- ggplot(data = as.data.frame(sft_thresh),
       aes (x = sft_thresh$fitIndices[,1],
     y = sft_thresh$fitIndices[,5],
     label = trial_powers)) + 
  geom_label()+
  
  # the sign expression returns the vector of signs of estimated slopes, multiplied by the correlation coefficient.

     labs( x="Soft Threshold (power)",
     y="Mean Connectivity", title = "Mean Connectivity", tag = "B")+
  theme_classic()


plot_grid(p1,p2,ncol=2)

```

## Topological Overlap Measure

The next steps is to caliberate the **Topological Overlap Measure (TOM)**. It is premised over the notion of shared relationships between two clusters. 

```{r TOM}

## Now we generate adjacency values and TOM similarity matrices on the basis of selected softpower.

softPower = 7;

#calclute the adjacency matrix
adj= adjacency(exprdata,
               type = "unsigned", 
               power = softPower);

#turn adjacency matrix into topological overlap to minimize the effects of noise and spurious associations
TOM = TOMsimilarityFromExpr(exprdata,
                            networkType = "unsigned", 
                            TOMType = "unsigned", 
                            power = softPower);

colnames(TOM) = rownames(TOM) = names(exprdata) # same rows as columns, therefore adjacency.
dissTOM=1-TOM # "distance/ dissimilarity = 1 - similarity"

```

## Module Detection

```{r clustering samples}

#hierarchical clustering of the genes based on the TOM dissimilarity measure

library(flashClust)
gene_tree = flashClust(as.dist(dissTOM),
                      method="average")

#plot the resulting clustering tree (dendrogram)
plot(gene_tree, 
     xlab="", 
     sub="",
     cex=0.3)

sample_tree <- as.dendrogram(hclust(dist(exprdata), method = "average"))

dplot <- ggdendrogram(data= sample_tree, rotate = FALSE)+
  theme_dendro()+
  ggtitle("Sample clustering to detect outliers")+
  theme(plot.title = element_text(hjust = 0.5))+
  xlab("Samples")+
  geom_label()

print(dplot)
```
```{r Tuning parameters for modules}

## Set the minimum module size

min_mod_size = 20

## Module identification using dynamic tree cut

dynamic_mods = cutreeDynamic(dendro = gene_tree,  
                             method="tree", 
                             minClusterSize = min_mod_size)

dynamic_mods1 = cutreeDynamic(dendro = gene_tree, 
                              distM = dissTOM, 
                              method="hybrid", 
                              deepSplit = 2, 
                              pamRespectsDendro = FALSE, 
                              minClusterSize = min_mod_size)


## we could highlight module labels and the size(number of nodes) of each module. Label 0 is reserved for unassigned genes that are in essence independent nodes.

table(dynamic_mods)
table(dynamic_mods1)

```

It is always better to have a enhanced, color-coded visualization of modules for precise interpretation.

```{r Modules in colors}

## we can conveniently assign colors to these labels. "Grey" is assigned to label 0.

dynamic_mod_colors = labels2colors(dynamic_mods)
table(dynamic_mod_colors)


dynamic_mod_colors1 = labels2colors(dynamic_mods1)
table(dynamic_mod_colors1)

  
```

We can also choose to discard the independent nodes (since they do not belong to any network) and just map the coeherent modules.

```{r Modules only}

## discard the unassigned genes, and focus on the rest

rest_genes= (dynamic_mod_colors != "grey")
dissTOM1 = 1 - TOMsimilarityFromExpr(exprdata[,rest_genes], 
                                  power = softPower)

## Let's plot the tree now.

colnames(dissTOM1) = rownames(dissTOM1) = names(exprdata[rest_genes])
hier_tree = flashClust(as.dist(dissTOM1), method="average" )
plotDendroAndColors(hier_tree, 
                    dynamic_mod_colors[rest_genes], 
                    "Dynamic Tree Cut", 
                    dendroLabels = FALSE, 
                    hang = 0.03, 
                    addGuide = TRUE, 
                    guideHang = 0.05, 
                    main = "Gene dendrogram and module colors")


```


```{r}

## set the diagonal of the dissimilarity to NA 

diag(dissTOM1) = NA;

## Visualize the TOM plot. Raise the dissimilarity matrix to the power of 4 to bring out the module structure

sizeGrWindow(7,7)
TOMplot(dissTOM1, hier_tree, as.character(dynamic_mod_colors[rest_genes]))

```

Let us extract the modules now.

```{r Module Extraction}

module_colors= setdiff(unique(dynamic_mod_colors), "grey")
for (color in module_colors){
    module = names(exprdata)[which(dynamic_mod_colors == color)]
    write.table(module, 
                paste0("module_",color, ".txt"), 
                sep="\t", 
                row.names=FALSE, 
                col.names=FALSE,
                quote=FALSE)
}

```

Look at expression patterns of these genes, as they're clustered.

```{r}

module.order <- unlist(tapply(1:ncol(exprdata),
                              as.factor(dynamic_mod_colors),
                              I))
m <- t(t(exprdata[,module.order])/apply(exprdata[,module.order],2,max))

heatmap (t(m),
         zlim = c(0,1),
         col = gray.colors(100),
         Rowv = NA,
         Colv = NA,
         labRow = NA,
         scale = "none",
         RowSideColors = dynamic_mod_colors[module.order])

```

We can now look at the module gene listings and try to interpret their functions .. for instance using http://amigo.geneontology.org/rte

Quantify module similarity by eigengene correlation. Eigengenes: Module representatives.

```{r}

MEList <-  moduleEigengenes(exprdata, colors = dynamic_mod_colors)
MEs <- MEList$eigengenes
plotEigengeneNetworks(MEs, "", 
                      marDendro = c(0,4,1,2), 
                      marHeatmap = c(3,4,1,2))
```

